What's in a name? Automated topic naming of software maintenance activities

### 2010-08-09 icse/icse2011.tex
ICSE 2011 file and figures.
Signed-off-by: Neil Ernst <neil@neilernst.net>

%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode
%\documentclass[10pt, conference, compsocconf]{IEEEtran}
\documentclass[letterpaper]{acm_proc_article-sp}
%\usepackage{fontspec}
%\usepackage{xunicode}
%\usepackage{xltxtra}
\usepackage{booktabs}
%\defaultfontfeatures{Mapping=tex-text}
%\setmainfont{Times New Roman}
%\usepackage{amsthm}
\usepackage{color}

@@

\usepackage{comment}
\newtheorem{mydef}{definition}
\usepackage{graphicx}
\newcommand{\slbl}[1]{\textbf{\textsf{#1}}}
\usepackage{multirow}%,multicolumn}
\usepackage{slashbox}
\usepackage{wrapfig}
\usepackage{booktabs, colortbl}
\usepackage{amsmath}
%\usepackage{amsmath,amssymb}
%\usepackage{algpseudocode}
%\usepackage[pdfstartview=FitH,xetex,bookmarksopenlevel=3,bookmarks=true]{hyperref}
% pdfpagescrop={92 112 523 778},
\usepackage[pdfstartview=FitH,bookmarksopenlevel=3,bookmarks=true]{hyperref} %pdfpagescrop={92 112 523 778},
%\usepackage{algorithm}%,algorithmic-fix}
\usepackage{url}
\begin{document}
\numberofauthors{5}
\author{
\alignauthor 
Abram Hindle\\
   \affaddr{David Cheriton School of Computer Science}\\
   \affaddr{University of Waterloo}\\
   \affaddr{Waterloo, Ontario, CANADA}\\
   \email{ahindle@uwaterloo.ca}
\alignauthor  Neil A. Ernst\\
   \affaddr{Department of Computer Science}\\
   \affaddr{University of Toronto}\\
   \affaddr{Toronto, Ontario, CANADA}\\
   \email{nernst@cs.toronto.edu}
\and
\alignauthor Michael W. Godfrey\\
   \affaddr{David Cheriton School of Computer Science}\\
   \affaddr{University of Waterloo}\\
   \affaddr{Waterloo, Ontario, CANADA}\\
   \email{migod@uwaterloo.ca}
\alignauthor Richard C. Holt \\
   \affaddr{David Cheriton School of Computer Science}\\
   \affaddr{University of Waterloo}\\
   \affaddr{Waterloo, Ontario, CANADA}\\
   \email{holt@uwaterloo.ca}
\alignauthor John Mylopoulos \\
   \affaddr{Department of Computer Science}\\
   \affaddr{University of Toronto}\\
   \affaddr{Toronto, Ontario, CANADA}\\
   \email{jm@cs.toronto.edu}}

\title{What's in a name? On the automated topic naming of software maintenance activities}

@@

%There has not been any   investigation into the latent topics these %classifiers represent.

%  Within the field of mining software repositories, many approaches,   such as topic modeling and concept location, rely on the automated   application of machine learning algorithms to corpora.  
%This kind of   approach typically makes no use of domain-specific knowledge to   improve the results or put them into a context that a developer can   relate to.  
%While too much specificity can produce non-generalizable   results, too little produces broad learners that do not provide much   immediately useful detail.  
%In this paper, we investigate what   topics the latent topic extractors produce, and examine whether   such topics can be generalized across software products.  
%We focus   on non-functional requirements related to software quality as a   potential generalization, since there is some shared belief that   these qualities apply broadly across many software systems and their   development artifacts.  
%In our approach, we name topics based on a   static ontology of non-functional requirements and software   qualities; in essence, this adds context to general-purpose   ontologies such as WordNet.  
%In this paper, we discuss how we use   this software development domain knowledge to better annotate topics   of development that are extracted from the change log messages in   configuration management systems such as CVS. 
% Since our taxonomy is   global in scope, our results allow us to compare the relative   importance of particular qualities across different projects. 
%Within the topics extracted from MySQL and MaxDB we found that   non-functional requirements were common and often were trends across   time. 
%Our automated techniques can label topics without much computational overhead.

\begin{comment}
 Within the field of mining software repositories, 
to have meaningful discussions about how software development is progressing, we must be able to devise appropriate labels for work topics. We can then categorize development artifacts as belonging to one or more of these topics. 
% Other approaches
In order to do this, approaches such as topic modeling and concept location rely on the automated application of machine learning algorithms to development artifacts. 
%limitations
However, these approaches neither use domain-specific knowledge to improve results, nor contextualize those results for developers.  While too much specificity can produce non-generalizable results, too little produces broad learners that do not provide much immediately useful detail. 
%introduce idea
This paper implements \emph{labelled topic extraction}, in which topics are extracted from commit comments and given labels relating to a cross-project taxonomy. We focus on non-functional requirements related to software quality as a potential generalization, since there is some shared belief that these qualities apply broadly across many software systems and their development artifacts.  
%In our approach, we name topics based on a static ontology of non-functional requirements and software qualities; in essence, this adds context to general-purpose ontologies such as WordNet.  
%elaborate
%We discuss how we use this software development domain knowledge to better annotate topics of development that are extracted from the change log messages in configuration management systems such as CVS. 
We evaluated our approach with an experimental study on two large-scale database projects, MySQL and MaxDB. We extracted topics from their version control commit logs comments. Our results were generalizable across the two projects, showing that non-functional requirements were commonly discussed, and we identified topic trends over time. 	
%Our automated techniques can label topics with little computational overhead.  
%sum up
% Since our taxonomy is global in scope, our results allow us to compare the relative importance of particular qualities across different projects.
Our labelled topic extraction technique allowed us to devise appropriate, context-sensitive labels across these two projects, providing insight into software development activities.
\end{comment}

\begin{abstract}
 Within the field of mining software repositories, many approaches,   such as topic modeling and concept location, rely on the automated   application of machine learning algorithms to corpora.  
Unfortunately the output of these tools is often difficult to distinguish and interpret as they are often so abstract.
Thus to have a meaningful discussion about the topics of software development, we must be able to devise appropriate labels for extracted topics. 
However, these approaches neither use domain-specific knowledge to improve results, nor contextualize those results for developers.   While too much specificity can produce non-generalizable results, too little produces broad learners that do not provide much immediately useful detail. 
This paper implements \emph{labelled topic extraction}, in which topics are extracted from commit comments and given labels relating to a cross-project taxonomy. We focus on non-functional requirements related to software quality as a potential generalization, since there is some shared belief that these qualities apply broadly across many software systems and their development artifacts.  
We evaluated our approach with an experimental study on two large-scale database projects, MySQL and MaxDB. We extracted topics using Latent Dirichlet Allocation (LDA) from the commit log comments of their version control systems (CVS and BitKeeper). Our results were generalizable across the two projects, showing that non-functional requirements were commonly discussed, and we identified topic trends over time. 	
Our labelled topic extraction technique allowed us to devise appropriate, context-sensitive labels across these two projects, providing insight into software development activities.
\end{abstract}

%\vspace*{4em} % push intro down

@@

\section{Introduction}
\begin{quote}
	\emph{What's in a name? that which we call a rose\\
	By any other name would smell as sweet;}
	\begin{flushright} -- Romeo and Juliet, II:ii\end{flushright}
\end{quote}

Few would argue that software development is a bed of roses.  To most of us, the important properties of a rose are sensual in nature, and concern appearance, odour, and feel.  While horticulturalists and botanists may see more semantic depth than this, one rose is often just as pleasing to us as another.  Software development artifacts, on the other hand, are abstract and intangible.  They are often unnamed and ephemeral --- even more so than roses --- and maybe partially derived from other unnamed and ephemeral artifacts.  Yet to have meaningful discussions about how software development is progressing, we must be able to devise appropriate labels for our development topics and be able to categorize the development artifacts as belonging to one or more of these topics.

Topics arise from the wide variety of issues which occur during a software project's life-cycle. These topics can relate to, among others, the problem domain, the processes and tools used, or the development artifacts themselves.  The set of development topics for a given project can sometimes be extracted automatically by analyzing artifacts within software repositories, such as change-log comments that developers create when committing revisions to the project's source control system.  Our previous work dealt with topic trends, which are topics that recur over time~\cite{Hindle09ICSM}. We observed that topic trends were often non-functional requirements.

Topics in this paper and the previous paper are word bags or word distributions. These word distributions are found via Latent Dirichlet Allocation~\cite{Blei2003}, which finds independent word distributions shared among documents (change-log comments). The unfortunate aspect of topics that are word distributions is that they lack the tangibility of a rose. They do not self identify with their tangible properties. Topics have to be identified by interpreting the prevalent words in the word distribution and by inspecting related documents. This is impractical when one has to handle more than one hundred different topics. It would be nice if we could have automatic assistance to determine what the topic is about. This is the power of labelling and naming.

A topic --- that is, a word distribution in our world --- needs a suggestive name to succinctly convey its intent, and make it easy to use in discussions and analyses about the development of the system. Our previous experience leads us to believe that automated topic naming is currently infeasible in the general case. We have therefore decided to focus on topic labels from the sub-domain of non-functional requirements related to software quality.

Traditionally topic extraction has required manual annotation to derive domain-relevant labels. This paper implements \emph{labelled topic extraction}, in which topics are extracted and given labels relating to a cross-project taxonomy. This is an ontological approach where we attempt to relate networks of words to labels and then search for these terms within our topics. We also compare this approach to machine learners.

Our contributions include:
\begin{itemize}
\item  We introduce labelled topic extraction, and demonstrate its usefulness
  compared to other approaches.

\item We show that these labels with their topics can be learnt and used to classify other data-sets.
\item We present visualizations of named topics and their trends over time to aid communication and analysis.
\item We use an exploratory case study of
  several open source database systems to show how named topics can be compared between projects.
\end{itemize}

We first introduce some important terminology for our work. We then describe our methodology, including our data-sets, then highlight our results. We conclude with a look at related work and possible improvements.

\shrink
\section{Background}

We provide a brief overview of software repository mining and information retrieval.
This work is related to the mining software repositories (MSR)~\cite{msr} research community as it deals expressly with analyzing a project's source control repository, and the messages associated with revisions therein.

\shrink
\subsection{Definitions}
We will use the following terminology in the rest of the paper.
A \emph{message} is a block of text written by a developer. In this
paper, messages will be the CVS and BitKeeper commit log comments made
when a developer commits changes to files in a repository. A \emph{word
  distribution} is the summary of a message by word count. Each word
distribution will be over the words in all messages. However, most
words will not appear in each message. A word distribution is effectively
a word count divided by the message size. A \emph{topic} is a word
distribution, i.e., a set of words that form a word distribution that is
unique and independent within the set of documents in our total
corpus. One could think of a topic as a distribution of the centroid
of a group of messages. In this paper we often summarize topics by the
top ten most frequent words of their word distribution.  A \emph{trend}
is one or more similar topics that recur over time.  Trends are
particularly important, as they indicate long-lived and recurring
topics that may provide key insights into the development of the
project. A \emph{label} is part of a title we attach to a topic, whether manually or automatically.

% area under the ROC 
% check
\emph{Area of ROC Curve} is the area under the Receiver Operating Characteristic (\emph{ROC}) curve, sometimes referred to as \emph{AUC}. ROC values reflect a score, similar to school letter-grades (A is 0.9, C is 0.6), reflecting how well a particular learner performed for the given data. A ROC result of 0.5 would be equivalent to a random learner (randomly classifying data). ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) versus the false positive rate (1 - specificity). A perfect learner has a ROC value of 1.0, reflecting perfect recall and precision.

\shrink
\subsection{Topic and Concept Extraction}
Topic extraction, sometimes called concept extraction, uses tools such as Latent Dirichlet Allocation (LDA)~\cite{Blei2003} and Latent Semantic Indexing (LSI) to extract
independent word distributions (topics) from
	documents (commit log comments). Many researchers~\cite{marcus04wcre,Poshyvanyk2007,lukins08wcre,Linstead2007} have applied tools like LSI and LDA to mining software repositories, in particular analyzing source code, bugs or developer comments.
\begin{comment}
Ideally these extracted topics 
correlate with actual development topics that the developers discussed
during the development of the software project. For example, our technique might identify a collection of words: ``list next has iterator add'', which we then identify (manually) as concerning the topics of collections and the Iterator pattern.
\end{comment}

Typically a topic analysis tool like LDA will try to find $N$ independent word distributions found within the word distributions of all the messages. Linear combinations of these $N$ word distributions are then meant to be able to recreate the word distributions of all of the underlying messages. These $N$ word distributions effectively form topics: cross cutting collections of words relevant to one or more documents. Our problem is that these topics are not easy to interpret, as the underlying pattern is not clear. We feel that automatic labelling or naming of these topics would be helpful with respect to interpreting the subject of a topic. LDA extracts topics in an unsupervised manner; the algorithm relies solely on the source data with no human intervention.

In topic analysis a single document, such as a commit message, can be related to multiple topics. Representing documents as a mixture of topics maps well to source code repository commits, which often have more than one purpose~\cite{Hindle09ICSM}.  A topic represents both a word distribution and a group of commit log comments that are related to each other by their content.  In this paper a topic is a set of tokens extracted from commit messages found within a project's source control system (SCS).

\begin{comment}
In this paper we use LDA to identify topics.  LDA works by creating topic models for each period. A topic model is a statistical summary of the important subjects for that data-set.  The input to the algorithm is a set of documents, with contents treated as a bag of words, and a parameter to set the number of topics to generate. The algorithm returns a word-document matrix relating the occurrence frequency of a 
 The idea is similar to other IR approaches like TF/IDF (term frequency/inverse document frequency), which find uncommon words that are not found in other word-lists. These approaches are unsupervised in that they rely solely on the source data with no human intervention.
\end{comment}

One issue that arises with use of unsupervised techniques is how to label the topics. While the topic models themselves are generated automatically, what to make of them is less clear. For example, in our previous work~\cite{Hindle09ICSM}, as well as in Baldi et al.~\cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a keyword accordingly. E.g., for the word list \emph{``listener change remove add fire''}, Baldi et al. assign the keyword \emph{event-handling}. The labels are reasonable enough, but still require an expert in the field to determine them. Our technique is automated, because we match keywords from WordNet~\cite{Fellbaum1998} to words in the topic model.

\shrink
\subsection{Supervised learning}
While unsupervised techniques (LSI and LDA are both unsupervised) are appealing in their lack of human intervention, and thus lower effort, supervised learners have the advantage of domain knowledge which typically means improved results. In supervised learning, the data is divided into slices. One slice is manually annotated by the domain expert, and the classifications he/she determines are applied to the remaining slices. In this paper, we employ the WEKA~\cite{weka09} and Mulan~\cite{mulan} machine learning frameworks in order to test supervised learning.

\shrink 
\section{Methodology}

\begin{comment}
% To gather the data, we first extract the commit log comments from a
% project's SCS repository. We filter out stop words and produce word distributions from these messages. These distributions are bucketed
% into 30-day non-overlapping windows. Each window is subject to LDA analysis. %Figure \ref{fig:commits}
% depicts the general process for processing and analyzing the commit messages.
\end{comment}

@@

\shrink
\subsection{Generating the data}
For each project, we used source control commit comments, the messages that programmers write when they commit revisions to a source control repository. We leveraged the data that we gathered in \cite{Hindle09ICSM} for this work. An example of a typical commit message is: \textit{``history annotate diffs bug fixed (if mysql\_real\_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course)''}. We extracted these messages and indexed them by creation time. We summarized each message as a word distribution but removed stop-words such as common English words like \emph{the} and \emph{at}.

From that data-set, we created an XML file which separated commits into
monthly windows. This size of period is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze. We applied Blei's LDA implementation~\cite{Blei2003} against the word distributions of these commits, and generated lists of topics per period. 
%A topic extracted in the first period Period 1, the most important %topic might have been ``security", with ``sql'' second, down on to %less relevant topics. 
We arbitrarily set the number of topics to generate to $20$, because past experimentation showed that fewer topics might aggregate multiple unique topics while any more topics seemed to dilute the results and create indistinct topics. As well, more than $20$ topics quickly became infeasible for inspection and it was difficult to discern the difference in topics.

\shrink
\subsection{Associating labels}
\begin{comment}
% We would like to `name' our topic models for each period. In this way, we can see how certain words -- subjects -- occur more or less often with each period.
\end{comment}

@@

To construct these expanded word-lists, we used WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words, including meronymy and synonymy. We then added Boehm’s 1976 software quality model \cite{Boehm+:1976:ICSE}, and classified his eleven ‘ilities’ into their respective ISO9126 qualities. We did the same for the quality model produced by McCall et al. \cite{mccall1977}. Finally, we analyzed two mailing lists from the KDE project to enhance the specificity of the sets. We selected KDE-Usability, which focuses on usability discussions for KDE as a whole; and KDE-Konqueror, a mailing list about a long-lived web browser project. For each high-level quality in ISO9126, we first searched for our existing labels; we then randomly sampled twenty-five mail messages that were relevant to that quality, and selected co-occurring terms relevant to that quality. For example, we add the term ``performance'' to the synonyms for \emph{efficiency}, since this term occurs in most KDE mail messages that discuss efficiency.
%exp2

@@

\begin{table*}
	\centering
\begin{tabular}{c|p{9cm}}
\toprule
\textbf{Label} & \textbf{Related terms} \\
\midrule
\emph{Maintainability} &
testability changeability analyzability stability maintain maintainable modularity modifiability understandability + interdependent dependency encapsulation decentralized modular\\ \hline
\emph{Functionality} &
security compliance accuracy interoperability suitability functional practicality functionality + compliant exploit certificate secured “buffer overflow” policy malicious trustworthy vulnerable vulnerability accurate secure vulnerability correctness accuracy\\ \hline
\emph{Portability} &
conformance adaptability replaceability installability portable movableness movability portability + specification migration standardized l10n localization i18n internationalization documentation interoperability transferability\\ \hline
\emph{Efficiency} &
“resource behaviour” “time behaviour” efficient efficiency + performance profiled optimize sluggish factor penalty slower faster slow fast optimization\\ \hline
\emph{Usability} &
operability understandability learnability useable usable serviceable usefulness utility useableness usableness serviceableness serviceability usability + gui accessibility menu configure convention standard feature focus ui mouse icons ugly dialog guidelines click default human convention friendly user screen interface flexibility\\ \hline
\emph{Reliability} &
“fault tolerance” recoverability maturity reliable dependable responsibleness responsibility reliableness reliability dependableness dependability + resilience integrity stability stable crash bug fails redundancy error failure\\ 
\bottomrule
\end{tabular}
	\caption{Qualities and associated signifiers – WordNet version (\textsf{exp2})}
	\label{tbl:wnsig}

@@

\shrink
\subsection{Supervised learning}

@@

\shrink
\section{Observations and evaluation}

\shrink
\subsection{Word list similarity}

@@

\shrink
\subsection{Analysis of the unsupervised labelling}

@@

Based on the labels, and our manual topic labelling, we compared the results of the unsupervised word matching approach. For each quality we tried to assess whether the manual tag matched the unsupervised label assigned. Table \ref{tbl:maxdb-unsup-results} shows our results for MaxDB and MySQL. In general results are poor. Using the F-Measure, the weighted average of precision and recall, where 1 is perfect, our best results are 0.6, a few at 0.4, and most around 0.2. We achieved similar results using the Matthew's correlation coefficient (used to measure efficacy where classes are of different sizes) and ROC.

\begin{table*}%[ht]
	\centering

>>>>

 \documentclass[]{sig-alternate}
 \usepackage{booktabs}
% \usepackage[colorinlistoftodos]{todonotes}
 \usepackage{color}

@@

\usepackage{comment}
\newtheorem{mydef}{definition}
\usepackage{graphicx}
\newcommand{\slbl}[1]{\textbf{\textsf{#1}}}
\usepackage{multirow}%,multicolumn}
\usepackage{slashbox}
\usepackage{wrapfig}
\usepackage{booktabs, colortbl}
\usepackage{amsmath}
\usepackage[pdfstartview=FitH,bookmarksopenlevel=3,bookmarks=true]{hyperref} %pdfpagescrop={92 112 523 778},

\usepackage{url}
\begin{document}
\numberofauthors{5}
\author{
\alignauthor 
Abram Hindle\\
   \affaddr{David Cheriton School of Computer Science}\\
   \affaddr{University of Waterloo}\\
   \affaddr{Waterloo, Ontario, CANADA}\\
   \email{ahindle@uwaterloo.ca}
\alignauthor  Neil A. Ernst\\
   \affaddr{Department of Computer Science}\\
   \affaddr{University of Toronto}\\
   \affaddr{Toronto, Ontario, CANADA}\\
   \email{nernst@cs.toronto.edu}
\and
\alignauthor Michael W. Godfrey\\
   \affaddr{David Cheriton School of Computer Science}\\
   \affaddr{University of Waterloo}\\
   \affaddr{Waterloo, Ontario, CANADA}\\
   \email{migod@uwaterloo.ca}
\alignauthor Richard C. Holt \\
   \affaddr{David Cheriton School of Computer Science}\\
   \affaddr{University of Waterloo}\\
   \affaddr{Waterloo, Ontario, CANADA}\\
   \email{holt@uwaterloo.ca}
\alignauthor John Mylopoulos \\
   \affaddr{Department of Computer Science}\\
   \affaddr{University of Toronto}\\
   \affaddr{Toronto, Ontario, CANADA}\\
   \email{jm@cs.toronto.edu}
}

\title{What's in a name? Automated topic naming of software maintenance activities}

@@

\begin{abstract}
%  Within the field of mining software repositories, many approaches,   such as topic modeling and concept location, rely on the automated   application of machine learning algorithms to corpora.  
% Unfortunately the output of these tools is often difficult to distinguish and interpret as they are often so abstract.
% Thus to have a meaningful discussion about the topics of software development, we must be able to devise appropriate labels for extracted topics. 
% However, these approaches neither use domain-specific knowledge to improve results, nor contextualize those results for developers.   While too much specificity can produce non-generalizable results, too little produces broad learners that do not provide much immediately useful detail. 
% This paper implements \emph{labelled topic extraction}, in which topics are extracted from commit comments and given labels relating to a cross-project taxonomy. We focus on non-functional requirements related to software quality as a potential generalization, since there is some shared belief that these qualities apply broadly across many software systems and their development artifacts.  
% We evaluated our approach with an experimental study on two large-scale database projects, MySQL and MaxDB. We extracted topics using Latent Dirichlet Allocation (LDA) from the commit log comments of their version control systems (CVS and BitKeeper). Our results were generalizable across the two projects, showing that non-functional requirements were commonly discussed, and we identified topic trends over time. 	
% Our labelled topic extraction technique allowed us to devise appropriate, context-sensitive labels across these two projects, providing insight into software development activities.
\end{abstract}

@@

\section{Introduction}
The act of developing software generates many artifacts, including source code, bug reports, commit messages, and test cases. It does not take long for a software project to grow to a point where keeping track of these artifacts exceeds easy comprehension. For these projects, developers interact with abstractions of these artifacts, rather than the artifact itself. Such abstractions include project dashboards~\cite{}, tags (Peggy/Christoph), and complexity metrics (\cite{mccabe}).

Developers do more than merely observe these abstractions, however \cite{}. They must also use them to assess progress and work with colleagues. To have meaningful discussions about how software development is progressing, we must be able to devise appropriate labels for our development artifacts. Such labels can be used to compare progress or understand how well a project is meeting requirements. These labels then become the topics for development conversations (such as the stream of commentary on a bug report). One way to identify topics is to use the artifacts themselves to derive them.

Deriving such topics is difficult. Topics have to be identified by interpreting the prevalent words in a word distribution and by inspecting related documents. This is impractical when one has to handle more than one hundred different topics. To scale to meaningfully sized projects, we require automatic assistance to determine the topic label.

In our previous work we dealt with topic trends, which are topics that recur over time~\cite{Hindle09ICSM}. We observed that topic trends were often non-functional requirements (NFRs). NFRs have the property of being cross-domain and widely applicable. In this sense, they are useful abstractions for developer conversations about different software projects.

In this paper, we describe \textbf{Nomen}, which is a tool for \emph{labelled topic extraction}. \textbf{Nomen} addresses two gaps in the topic mining literature:
\begin{enumerate}
  \item Topic mining of software has been limited to one project at a time. This is because traditional topic mining techniques are specific to a particular dataset. \textbf{Nomen} allows for comparisons \textit{between} projects. 
  \item Topic modeling creates bags of words that require interpretation by the user to assign meaning. Like (1), this means
that it is difficult to discuss results independent of the project context. Our technique automatically assigns labels agnostic of project.
\end{enumerate}

This paper begins by introducing labelled topic extraction, demonstrating its usefulness compared to other approaches. We show that these labels with their topics can be learned and used to classify other data-sets. We then present visualizations of named topics and their trends over time to aid communication and analysis. We use an exploratory case study of two open source database systems to show how named topics can be compared between projects.

% \section{Background}

% We provide a brief overview of software repository mining and information retrieval.
% This work is related to the mining software repositories (MSR)~\cite{msr} research community as it deals expressly with analyzing a project's source control repository, and the messages associated with revisions therein.

% \subsection{Definitions}
% We will use the following terminology in the rest of the paper.
% A \emph{message} is a block of text written by a developer. In this
% paper, messages will be the CVS and BitKeeper commit log comments made
% when a developer commits changes to files in a repository. A \emph{word
%   distribution} is the summary of a message by word count. Each word
% distribution will be over the words in all messages. However, most
% words will not appear in each message. A word distribution is effectively
% a word count divided by the message size. A \emph{topic} is a word
% distribution, i.e., a set of words that form a word distribution that is
% unique and independent within the set of documents in our total
% corpus. One could think of a topic as a distribution of the centroid
% of a group of messages. In this paper we often summarize topics by the
% top ten most frequent words of their word distribution.  A \emph{trend}
% is one or more similar topics that recur over time.  Trends are
% particularly important, as they indicate long-lived and recurring
% topics that may provide key insights into the development of the
% project. A \emph{label} is part of a title we attach to a topic, whether manually or automatically.

% % area under the ROC 
% % check
% \emph{Area of ROC Curve} is the area under the Receiver Operating Characteristic (\emph{ROC}) curve, sometimes referred to as \emph{AUC}. ROC values reflect a score, similar to school letter-grades (A is 0.9, C is 0.6), reflecting how well a particular learner performed for the given data. A ROC result of 0.5 would be equivalent to a random learner (randomly classifying data). ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) versus the false positive rate (1 - specificity). A perfect learner has a ROC value of 1.0, reflecting perfect recall and precision.

% \shrink
% \subsection{Topic and Concept Extraction}
% Topic extraction, sometimes called concept extraction, uses tools such as Latent Dirichlet Allocation (LDA)~\cite{Blei2003} and Latent Semantic Indexing (LSI) to extract
% independent word distributions (topics) from
% 	documents (commit log comments). Many researchers~\cite{marcus04wcre,Poshyvanyk2007,lukins08wcre,Linstead2007} have applied tools like LSI and LDA to mining software repositories, in particular analyzing source code, bugs or developer comments.
% \begin{comment}
% Ideally these extracted topics 
% correlate with actual development topics that the developers discussed
% during the development of the software project. For example, our technique might identify a collection of words: ``list next has iterator add'', which we then identify (manually) as concerning the topics of collections and the Iterator pattern.
% \end{comment}

% Typically a topic analysis tool like LDA will try to find $N$ independent word distributions found within the word distributions of all the messages. Linear combinations of these $N$ word distributions are then meant to be able to recreate the word distributions of all of the underlying messages. These $N$ word distributions effectively form topics: cross cutting collections of words relevant to one or more documents. Our problem is that these topics are not easy to interpret, as the underlying pattern is not clear. We feel that automatic labelling or naming of these topics would be helpful with respect to interpreting the subject of a topic. LDA extracts topics in an unsupervised manner; the algorithm relies solely on the source data with no human intervention.

% In topic analysis a single document, such as a commit message, can be related to multiple topics. Representing documents as a mixture of topics maps well to source code repository commits, which often have more than one purpose~\cite{Hindle09ICSM}.  A topic represents both a word distribution and a group of commit log comments that are related to each other by their content.  In this paper a topic is a set of tokens extracted from commit messages found within a project's source control system (SCS).

% One issue that arises with use of unsupervised techniques is how to label the topics. While the topic models themselves are generated automatically, what to make of them is less clear. For example, in our previous work~\cite{Hindle09ICSM}, as well as in Baldi et al.~\cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a keyword accordingly. E.g., for the word list \emph{``listener change remove add fire''}, Baldi et al. assign the keyword \emph{event-handling}. The labels are reasonable enough, but still require an expert in the field to determine them. Our technique is automated, because we match keywords from WordNet~\cite{Fellbaum1998} to words in the topic model.

% \shrink
% \subsection{Supervised learning}
% While unsupervised techniques (LSI and LDA are both unsupervised) are appealing in their lack of human intervention, and thus lower effort, supervised learners have the advantage of domain knowledge which typically means improved results. In supervised learning, the data is divided into slices. One slice is manually annotated by the domain expert, and the classifications he/she determines are applied to the remaining slices. In this paper, we employ the WEKA~\cite{weka09} and Mulan~\cite{mulan} machine learning frameworks in order to test supervised learning.

\section{Methodology}

% To gather the data, we first extract the commit log comments from a  project's SCS repository. We filter out stop words and produce word distributions from these messages. These distributions are bucketed
% into 30-day non-overlapping windows (periods). Each window is subject to LDA analysis. We arbitrarily create 20 topics, vectors of important words according to LDA.

%Figure \ref{fig:commits}
% depicts the general process for processing and analyzing the commit messages.

Figure \ref{fig:process} gives an outline of the methodology followed.

\begin{figure*}
  \centering
 \includegraphics[width=.55\textwidth]{process-model}
 \caption{Research methodology process view.}
  \label{fig:process}
\end{figure*}

@@

\subsection{Generating the data}
For each project, we used source control commit comments, the messages that programmers write when they commit revisions to a source control repository. We leveraged the data that we gathered in \cite{Hindle09ICSM} for this work. An example of a typical commit message is: \textit{``history annotate diffs bug fixed (if mysql\_real\_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course)''}. We extracted these messages and indexed them by creation time. We summarized each message as a word distribution but removed stop-words such as \emph{the} and \emph{at}.

From that data-set, we created an XML file which separated commits into
monthly windows. This size of period is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze. We applied Blei's LDA implementation~\cite{Blei2003} against the word distributions of these commits, and generated lists of topics per period. 
%A topic extracted in the first period Period 1, the most important %topic might have been ``security", with ``sql'' second, down on to %less relevant topics. 
We arbitrarily set the number of topics to generate to $20$, because past experimentation, e.g. \cite{Hindle09ICSM}, showed that fewer topics might aggregate multiple unique topics while any more topics seemed to dilute the results and create indistinct topics. Furthermore, more than $20$ topics quickly became infeasible for inspection.

\subsection{Associating labels}
\begin{comment}
% We would like to `name' our topic models for each period. In this way, we can see how certain words -- subjects -- occur more or less often with each period.
\end{comment}

@@

To construct these expanded word-lists, we used WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words, including meronymy and synonymy. We then added Boehm's 1976 software quality model \cite{Boehm+:1976:ICSE}, and classified his eleven `ilities' into their respective ISO9126 qualities. We did the same for the quality model produced by McCall et al. \cite{mccall1977}. Finally, we analyzed two mailing lists from the KDE project to enhance the specificity of the sets. We selected KDE-Usability, which focuses on usability discussions for KDE as a whole; and KDE-Konqueror, a mailing list about a long-lived web browser project. For each high-level quality in ISO9126, we first searched for our existing labels; we then randomly sampled twenty-five mail messages that were relevant to that quality, and selected co-occurring terms relevant to that quality. For example, we add the term ``performance'' to the synonyms for \emph{efficiency}, since this term occurs in most KDE mail messages that discuss efficiency.
%exp2

@@

\begin{table*}
	\centering
\begin{tabular}{c|p{9cm}}
\toprule
\textbf{Label} & \textbf{Related terms} \\
\midrule
\emph{Maintainability} &
testability changeability analyzability stability maintain maintainable modularity modifiability understandability + interdependent dependency encapsulation decentralized modular\\ \hline
\emph{Functionality} &
security compliance accuracy interoperability suitability functional practicality functionality + compliant exploit certificate secured “buffer overflow” policy malicious trustworthy vulnerable vulnerability accurate secure vulnerability correctness accuracy\\ \hline
\emph{Portability} &
conformance adaptability replaceability installability portable movableness movability portability + specification migration standardized l10n localization i18n internationalization documentation interoperability transferability\\ \hline
\emph{Efficiency} &
“resource behaviour” “time behaviour” efficient efficiency + performance profiled optimize sluggish factor penalty slower faster slow fast optimization\\ \hline
\emph{Usability} &
operability understandability learnability useable usable serviceable usefulness utility useableness usableness serviceableness serviceability usability + gui accessibility menu configure convention standard feature focus ui mouse icons ugly dialog guidelines click default human convention friendly user screen interface flexibility\\ \hline
\emph{Reliability} &
“fault tolerance” recoverability maturity reliable dependable responsibleness responsibility reliableness reliability dependableness dependability + resilience integrity stability stable crash bug fails redundancy error failure\\ 
\bottomrule
\end{tabular}
	\caption{Qualities and associated signifiers -- WordNet version (\textsf{exp2})}
	\label{tbl:wnsig}

@@

\subsection{Supervised learning}

@@

\section{Observations and evaluation}

\subsection{Word list similarity}

@@

\subsection{Analysis of the unsupervised labelling}

@@

Based on the labels, and our manual topic labelling, we compared the results of the unsupervised word matching approach. For each quality we tried to assess whether the manual tag matched the unsupervised label assigned. Figure \ref{fig:maxdb-unsup-results} shows our results for MaxDB and MySQL. In general results are poor. Total ROC was slightly better than 0.5, and usability seems relatively accurate.

\begin{figure*}
  \centering
 \includegraphics[width=.55\textwidth]{roc-unsup-pretty}
 \caption{ROC values for each NFR, per project and wordlist.}
  \label{fig:maxdb-unsup-results}
\end{figure*}

% Using the F-Measure, the weighted average of precision and recall, where 1 is perfect, our best results are 0.6, a few at 0.4, and most around 0.2. We achieved similar results using the Matthew's correlation coefficient (used to measure efficacy where classes are of different sizes) and ROC.

% \begin{table*}%[ht]
% 	\centering
